### 2014-11-18 icaps15-paper/paper.tex
ICAPS-15 paper sources

\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}

\newcommand{\tuple}[1]{{\langle #1\rangle}}
\newcommand{\triple}[1]{{\langle #1\rangle}}
\newcommand{\pair}[1]{{\langle #1\rangle}}

\newcommand{\Omit}[1]{}

\newcommand{\propername}[1]{\textsc{#1}}
\newcommand{\eqdef}{\stackrel{\hbox{\tiny{def}}}{=}}
%\newcommand{\IW}{{\textit{IW}}}
%\newcommand{\SIW}{{\textit{SIW}}}
%\newcommand{\ID}{{\textit{ID}}}
%\newcommand{\BRFS}{{\textit{BrFS}}}
%\newcommand{\BFS}{{\textit{\BFS}}}
%\newcommand{\UCT}{{\textit{UCT}}}
\newcommand{\IW}{{\textsc{IW}}}
\newcommand{\SIW}{{\textsct{SIW}}}
\newcommand{\ID}{{\textsc{ID}}}
\newcommand{\BRFS}{{\textsc{BrFS}}}
\newcommand{\BFS}{{\textsc{2BFS}}}
\newcommand{\UCT}{{\textsc{UCT}}}

\newcommand{\Asterix}{\propername{Asterix}}
\newcommand{\BeamRider}{\propername{Beam Rider}}
\newcommand{\Breakout}{\propername{Breakout}}
\newcommand{\Enduro}{\propername{Enduro}}
\newcommand{\Freeway}{\propername{Freeway}}
\newcommand{\Pong}{\propername{Pong}}
\newcommand{\Qbert}{\propername{Q*Bert}}
\newcommand{\Seaquest}{\propername{Seaquest}}
\newcommand{\SpaceInvaders}{\propername{Space Invaders}}
\newcommand{\Deepmind}{\propername{Deepmind}}

% AAAI (comment for ECAI)
% \newtheorem{theorem}{Theorem}

 \newtheorem{theorem}{Definition}
% \newtheorem{proposition}[theorem]{Proposition}
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
 \newtheorem{definition}[theorem]{Definition}

 \begin{document}
% The file aaai.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%
\pdfinfo{
/Title (Classical Planning Algorithms on the Atari Video Games: Preliminary Results)
/Author ( )}
% (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}

\setcounter{secnumdepth}{0}

\title{Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results}

\author{\textbf{ID:} 52 \ \ ; \ \ \textbf{Keywords:} classical planning, video games}

% \author{Nir Lipovetzky \\
%       The University of Melbourne \\
%    Melbourne, Australia\\
%    {\normalsize\url{@unimelb.edu.au}}
%    \And
%    Miquel Ramirez \\
%    RMIT University \\
%    Melbourne, Australia\\
%   {\normalsize\url{@rmit.edu.au}}
%    \And
%    Hector Geffner \\
%    ICREA \&  Universitat Pompeu Fabra \\
%    Barcelona, SPAIN \\
%    {\normalsize\url{@upf.edu}}\thanks{firstname.lastname}
% }

%AAAI Press\\
%Association for the Advancement of Artificial Intelligence\\
%2275 East Bayshore Road, Suite 160\\
% Palo Alto, California 94303\\}

\maketitle

%% Pending: randomization, discount factor, iw11 ..
%% iw2 multivalued expands up to 256^2 n^nodes, while iw2 boolean 16^2 n^2 nodes; can iw2 boolean beat iw1 multivalued?
%% Important. optimization, need Dikjstra por postprocessing .. see emails

\begin{abstract}
\begin{quote}
The Atari  2600 games supported in  the Arcade Learning Environment (Bellemare et. al. 2013)
all feature a known initial (RAM) state and actions that have deterministic effects. Classical planners, however,
cannot be used  for selecting actions  for  two reasons: first,  no  compact PDDL-model of  the games is given,
and  more importantly,  the action effects and goals are  not known \emph{a priori}. Moreover, in these games there is usually
no set of goals to be achieved but rewards to be collected. These features do not preclude the use of   classical  algorithms like
breadth-first search or Dijkstra's algorithm,  but    these methods are not effective over  large state spaces. We thus turn to a
different class of classical planning algorithms  introduced recently  that perform a \emph{structured exploration} of the state space; namely,
like breadth-first search and Dijkstra's  algorithm they are ``blind'' and hence  do not require prior knowledge of state transitions,
costs (rewards) or goals, and yet, like  heuristic search algorithms, they have been shown to be   effective for solving   problems over huge
state spaces. The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, \ldots,
$\IW(k)$ where  $\IW(i)$ is  a breadth-first search in which a state is  pruned when it is not the
first state in the search  to make true some set of $i$ atoms. 
The  empirical results suggest that the performance of \IW\ with the  $k$ parameter fixed to $1$, 
i.e., $\IW(1)$, approaches the performance of the state of the art represented by 
 \UCT, and that a simple best-first variation  of \IW\  that combines exploration 
and exploitation is even more competitive.

\end{quote}
\end{abstract}

\section{Introduction}

The Arcade Learning Environment (ALE)  provides a challenging  platform for evaluating general, domain-independent AI planners and
learners through a convenient interface to hundreds of Atari 2600 games \cite{bellemare:jair2013}.
Results have been reported so far for basic planning algorithms like breadth-first
search and  \UCT,  reinforcement learning algorithms, and evolutionary  methods \cite{bellemare:jair2013,deep-mind-atari,risto-texas:atari-games}.
The empirical results are impressive in some cases, yet a lot remains to be done, as no method  approaches   the
performance of  human players across a broad range of games.

While all  these games feature a known initial (RAM) state and actions that
have deterministic effects, the problem of selecting the next action
to be done cannot be addressed with state-of-the-art classical planners \cite{geffner:book}.
This is because there is  no compact PDDL-like encoding of the domain, and more
importantly,  the goal to be achieved in each game  is not given. Indeed, there are
often no goals  but rewards $r(a,s)$ that  depend on the action $a$ and
the state $s$ in  which  the action is performed, and these rewards are not known.\footnote{Actually, in the Atari games, the rewards $r(a,s)$ depend only
on the state $f(a,s)$ that results from doing action $a$ in the state $s$.}
Thus no model of the  goals  or the  rewards can be used to bias the search.

The action selection problem in the Atari games can be seen as a \emph{reinforcement learning} problem
\cite{sutton:book} over a deterministic MDP where the state transitions and rewards are not known, or alternatively,
as a  \emph{net-benefit planning problem} \cite{coles2012survey,keyder:jair09} with unknown state transitions  and  rewards. 
Namely, we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the  current  state $s_0$ that generates
a state sequence $s_0, s_1, \ldots, s_{m+1}$ with  maximum  total reward $\sum_{i=0}^{m} r(a_i,s_i)$,
where $m$ is a sufficiently large planning horizon, and the rewards $r(a_i,s_i)$ and the  state-transitions $s_{i+1}=f(a_i,s_i)$
are known only after  action $a_i$ is applied in the state $s_i$.

The presence of unknown transition and  rewards in  the Atari games does not preclude the use of blind-search
methods like breadth-first search, Dijkstra's algorithm \cite{dijkstra},  or learning methods  such as
LRTA* \cite{korf:lrta}, \UCT\ \cite{uct}, and Q-learning  \cite{sutton:book,bertsekas:neuro}. Indeed,
the net-benefit planning problem with unknown state transitions and rewards over a given planning
horizon, can be mapped into a standard \emph{shortest-path problem} which can be solved optimally by Dijkstra's algorithm.
For this, we just need to  map the unknown rewards $r(a,s)$ into positive (unknown) action costs $c(a,s) = C - r(a,s)$
where  $C$ is a large constant that exceeds the maximum possible reward. The fact that the state
transition and cost functions  $f(a,s)$ and $c(a,s)$ are not known a priori doesn't affect the applicability of
Dijkstra's algorithm, which requires the value of these functions precisely when  the action $a$
is applied in the state $s$.

The limitation of the   basic blind search  methods is that they are not effective  over large  spaces, neither
for solving problems off-line, nor for guiding  a lookahead search for solving problems on-line.
In this work, we thus turn to a recent class of  planning algorithms that combine the scope of blind search
methods with the performance of state-of-the-art classical planners: namely, like ``blind'' search algorithms
they do not require prior knowledge of state transitions, costs, or goals, and yet like heuristic algorithms
they manage to search large  state spaces effectively. The basic algorithm in this class is
called \IW\ for Iterated Width search \cite{nir:ecai2012}. \IW\ consists of a sequence of calls  $\IW(1)$, $\IW(2)$, .., $\IW(k)$, where
$\IW(i)$ is  a  standard breadth-first search where states are pruned right away when they fail to make true some new tuple (set)
of at most $i$ atoms. Namely,   $\IW(1)$ is a breadth-first search that keeps  a state only if  the state is   the first  one
in the search to make some atom  true; $\IW(2)$ keeps a state  only if the state is  the first one to  make  a pair of atoms true, and
so on.  Like plain breadth-first  and iterative deepening searches,  \IW\  is
complete, while searching the state space in a way that makes use of the
\emph{structure of states} given by the values of a finite set of state variables.
In the Atari games, the (RAM) state is given by a vector of 128 bytes, which we associate with 128 variables
$X_i$, $i=1, \ldots, 128$, each of which may take up to $256$ values $x_j$. A state $s$ makes an atom $X_i=x_{j}$
true when the value of the $i$-th byte in the state vector $s$ is $x_j$.

While a normal, complete breadth-first search runs in time that is exponential in the total number of variables $X_i$,
the procedure $\IW(k)$ runs in time that is exponential in the $k$ parameter, $1 \leq k \leq n$.
Elsewhere, it has been shown that  \IW\ exhibits   state-of-the-art performance on most existing planning benchmark domains
when goals are restricted to single atoms  where \IW\ provably runs in low polynomial time in the number of
variables \cite{nir:ecai2012}. For example, in a blocks world instance with any number of blocks and any initial
configuration,  the goal of  placing a given  block on top of another will be achievable in  quadratic time
in the number of blocks as any such instance can be shown to  have \emph{width} 2. This means  that $\IW(2)$ will
not only be complete then, but also optimal. This is also true for most other benchmark domains in classical planning
that turn out to  have small, bounded widths for \emph{any} instance as long as the goal is atomic.
For dealing with the benchmarks where goals are  not single atoms, e.g., where a given \emph{tower} of blocks needs to be built,
simple extensions of the basic \IW\ procedure have been developed  such as Serialized \IW, where \IW\ is
used to achieve the goals one at at time following a simple goal ordering  \cite{nir:ecai2012,nir:ecai2014}.
The extension of \IW\ for problems where there is no goal but an
additive reward measure  to be optimized, like in the Atari games, is direct.

%%  Each iteration IW(k) keeps track of the accumulated reward up to each state $s$ that has been
%% generated but  not pruned, as $R(s) = R(s')+ r(a,s')$ where $s'$ and $a$ are  the single parent state
%% and action $a$ that led to $s$ in IW(k) with $R(s)=0$ for the current state $s$. The best reward

% \Omit{
% The sequence of calls IW(1), IW(2), \ldots, IW(k) manages to combine the scope of ``blind'' search algorithms with the performance of
% heuristic-search algorithms by exploring states  in an  order dictated by the ``width''  parameter $k$;
% states that have width 1 are explored first in linear time, then states that have  width two are explored in quadratic time, and so on.
% In classical planning problems, if the goal has a small width (something common in the classical planning domains when the  goal is a single atom, as
% we have seen above), IW will find goal states  in low polynomial time. Similarly, in the reward-based setting, if  there are states of high reward
% that have small width, IW will find such states in low polynomial time, even if the state space is exponential in the number of variables.
% In this paper, we aim to test whether these ideas work for the Atari games too  when  the RAM state of the emulator
% is used  as the system state.
% }

The paper is organized as follows. We review the iterated width  algorithm and where it comes from,
look at  simple   variations of the algorithm that appear to be convenient  for the Atari
games, present and analyze the  experimental results, and discuss related and future work.

\section{Iterated Width}

The Iterated Width  (\IW) algorithm has been introduced as a classical planning algorithm
that takes a planning problem as an input, and computes an action sequence that solves the problem
as the output \cite{nir:ecai2012}. The algorithm however
applies to a broader range of problems. We will characterize such
problems  by means of  a finite and discrete set of states (the state space)
that correspond to vectors of size $n$. Namely, the
states are  \emph{structured} or \emph{factored} , and we take each of the locations in the vector
to represent a variable $X_i$, and the value at that vector location to
represent the value $x_j$ of variable $X_i$ in the state.
In addition to the state space, a problem is defined by an initial state $s_0$,
a set of actions applicable in each state, a
transition function $f$ such that  $s' = f(a,s)$ is
the state that results from applying action $a$  to the state $s$,
and rewards $r(a,s)$ represented by real numbers that result from  applying action $a$ in state $s$.
The transition and reward functions do not need to be known a priori, yet in that case, the
state and  reward that results from the application of an action in a state needs to be \emph{observable}.
The task is to compute an action sequence $a_0, \ldots, a_m$ for a large horizon $m$
that generates a state sequence $s_0, \ldots, s_{m+1}$ that  maximizes the accumulated
reward $\sum_{i=0}^{m} r(a_i,s_i)$, or that provides a good approximation.

\subsection{The Algorithm}

\IW \ consists of a sequence of calls  $\IW(i)$ for $i=0, 1, 2, \ldots$
over a problem $P$ until a termination condition is reached.
The procedure $\IW(i)$  is a plain forward-state \emph{breadth-first search} with just one change: right after a state
$s$ is generated, the state is  pruned if it doesn't pass a simple \emph{novelty test}.
More precisely,

\begin{itemize}
\item The \emph{novelty of a newly generate state $s$  in a  search algorithm}
is   $1$ if  $s$ is the first state generated in the search that makes true some  atom $X=x$,
else it is $2$ if  $s$ is the first state that makes a \emph{pair} of atoms $X=x$ and $Y=y$ true,
and so on.

\item  $\IW(i)$ is a breadth-first search that prunes newly generated states when their novelty measure
is greater than $i$.

\item $\IW$  calls $\IW(i)$ sequentially for $i=1,2,\ldots$ until a termination condition is reached,
returning then the best path found.
\end{itemize}

For classical planning, the termination condition is the achievement of the goal. In the on-line setting,
as in the Atari games, the termination condition is given by a time window or a maximum number of generated nodes.
The  best path found by \IW \ is the path  that has a  maximum accumulated reward.
The accumulated reward $R(s)$  of a state $s$ reached in an iteration of \IW \ is determined by the  unique parent state $s'$
and action $a$ leading to $s$ from $s'$ as $R(s) = R(s') + r(a,s')$. The best state is the state $s$ with maximum reward
$R(s)$ generated but not pruned by \IW, and the best path is the one that leads to the state $s$ from the current state.
The action selected to be done next is  the first action along  such a path.

%% Like in standard breadth-first search, each iteration IW(i)  prunes
%%% the  duplicate states as well.

\begin{table}[t!]
\begin{center}
%\resizebox{3in}{!}{
{\scriptsize
% \begin{tabular}{r@{}lrrrr@{}}
\begin{tabular}{llcccc}
%% table header
\hline\\[-1.3ex]
%\toprule
\# & \multicolumn{1}{c}{Domain} & \multicolumn{1}{c}{I} & \multicolumn{1}{c}{$\IW(1)$} & \multicolumn{1}{c}{$\IW(2)$} & \multicolumn{1}{c}{Neither} \\[.2ex]

\hline\\[-1.3ex]
%\hline
%% table body
1. & 8puzzle	&400	&55\%	&45\%	&0\%\\
2. & Barman	&232	&9\%	&0\%	&91\%\\
3. & Blocks World	&598	&26\%	&74\%	&0\%\\
4. &Cybersecure	&86	&65\%	&0\%	&35\%\\
% 5. &Depots	&189	&11\%	&66\%	&23\%\\
% 6. & Driver	&259	&45\%	&55\%	&0\%\\
% 7. & Elevators	&510	&0\%	&100\%	&0\%\\
% 8. & Ferry	&650	&36\%	&64\%	&0\%\\
% 9. & Floortile	&538	&96\%	&4\%	&0\%\\
% 10. & Freecell	&76	&8\%	&92\%	&0\%\\
% 11. & Grid	&19	&5\%	&84\%	&11\%\\
% 12. & Gripper	&1275	&0\%	&100\%	&0\%\\
% 13. & Logistics	&249	&18\%	&82\%	&0\%\\
% 14. & Miconic	&650	&0\%	&100\%	&0\%\\
% 15. & Mprime	&43	&5\%	&95\%	&0\%\\
% 16. & Mystery	&30	&7\%	&93\%	&0\%\\
% 17. & NoMystery	&210	&0\%	&100\%	&0\%\\
% 18. & OpenStacks	&630	&0\%	&0\%	&100\%\\
% 19. & OpenStacksIPC6	&1230	&5\%	&16\%	&79\%\\
\ldots & \ldots & \ldots & \ldots & \ldots & \\
20. & ParcPrinter	&975	&85\%	&15\%	&0\%\\
21. & Parking	&540	&77\%	&23\%	&0\%\\
22. & Pegsol	&964	&92\%	&8\%	&0\%\\
23. & Pipes-NonTan	&259	&44\%	&56\%	&0\%\\
24. & Pipes-Tan	&369	&59\%	&37\%	&3\%\\
25. & PSRsmall	&316	&92\%	&0\%	&8\%\\
26. & Rovers	&488	&47\%	&53\%	&0\%\\
27. & Satellite	&308	&11\%	&89\%	&0\%\\
28. & Scanalyzer	&624	&100\%	&0\%	&0\%\\
29. & Sokoban	&153	&37\%	&36\%	&27\%\\
30. & Storage	&240	&100\%	&0\%	&0\%\\
31. & Tidybot	&84	&12\%	&39\%	&49\%\\
32. & Tpp	&315	&0\%	&92\%	&8\%\\
33. & Transport	&330	&0\%	&100\%	&0\%\\
34. & Trucks	&345	&0\%	&100\%	&0\%\\
35. & Visitall	&21859	&100\%	&0\%	&0\%\\
36. & Woodworking	&1659	&100\%	&0\%	&0\%\\
37. & Zeno	&219	&21\%	&79\%	&0\%\\[.2ex]
%\hline
\hline\\[-1.3ex]
& Total/Avgs&37921	&37.0\%	&51.3\%	&11.7\%\\
\hline
\end{tabular}
}

\vskip .1cm

{\scriptsize
\begin{tabular}{@{}ccccc@{}}
\hline\\[-1.3ex]
\multicolumn{1}{c}{\# Instances}  &\multicolumn{1}{c}{\IW} & \multicolumn{1}{c}{\ID} & \multicolumn{1}{c}{\BRFS} &\multicolumn{1}{c}{\textit{GBFS +} $h_{add}$} \\[.2ex]
\hline\\[-1.3ex]
37921  &34627 &9010	&8762 &34849 \\
\hline
%\hline
\end{tabular}
}

\end{center}

\caption{\small
\emph{Top:} Number of classical planning instances per domain  and percentages solved by $\IW(1)$, $\IW(2)$, or neither.
Problems  obtained by splitting benchmarks  with $N$ atomic goals into $N$ problems with atomic goals.
\emph{Bottom:} Number of instances solved by \IW\ in comparison with  Iterative Deepening (\ID),
Breadth-First Search (\BRFS),   and Greedy Best First Search (\textit{GBFS}) guided by  additive heuristic.
Time and  memory outs  after 30 minutes or 2 GB. Table from \cite{nir:ecai2012}.
}
\label{table1}
\end{table}
%%%%
%%%

\subsection{Performance and Width}

\IW\ is a systematic and complete blind-search algorithm like breadth-first search (\BRFS) and iterative deepening (ID),
but   unlike these algorithms, it uses the factored representation of the states in terms of variables to structure the
search in a different way. This structured exploration  has proved to be very
effective over classical planning benchmark domains  when goals are  single atoms.\footnote{
Any conjunctive goal can be mapped into a single dummy atomic goal by adding an action that
achieves the dummy goal and that has the original conjunctive goal as a precondition.
Yet, this changes the definition of the domain.}
Table~\ref{table1} from \cite{nir:ecai2012} shows
the percentage of instances that are  solved by  the first iteration of \IW, i.e.  $\IW(1)$, by the second iteration
$\IW(2)$, and by neither one.  These are instances that  have been obtained from the existing benchmarks
by splitting problems with $N$ atomic goals, into $N$ problems with one atomic goal each. As the table shows,
$37\%$ of the 37921 instances are solved by $\IW(1)$ while $51.3\%$ are  solved by $\IW(2)$. Since $\IW(k)$
runs in time that is exponential in $k$, this mean that almost $90\%$ of the 37,921 instances are  solved in time that
is either linear or quadratic in the number of problem variables, which in these encoding are all \emph{boolean}.
Furthermore, when the performance of \IW\   is compared with  breadth-first search and iterative deepening,
on the one hand,  and  with a  Greedy Best First Search guided by the additive heuristic $h_{add}$
\cite{bonet:aij-hsp} on the other (this algorithm is similar
to the one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}),
it turns out that ``blind'' \IW\ solves as many problems as the informed search, 34,627 vs. 34,849, far ahead
of the other blind algorithms \BRFS\ and \ID\  that solve 9,010 and 8,762 problems each. This is shown in the bottom
part of  Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than a heuristic search algotrithm
\cite{nir:ecai2012}.

The min $k$ value for which $\IW(k)$ solves a problem  is bounded and small in most of
these instances. This is  actually no  accident and  has a \emph{theoretical explanation.} Lipovetzky and Geffner define
a structural parameter called the problem \emph{width} and show that for many of these domains,
\emph{any} solvable instance with atomic goals will  have a bounded and small width that is independent of the number
of variables and states in the problem. The min value $k$ for which the iteration  $\IW(k)$ solves
the problem cannot exceed the problem width, so  the algorithm \emph{\IW\ runs in time and space
that are exponential in the problem width.}

Formally, the \emph{width} $w(P)$ of a problem $P$ is $i$
iff $i$ is the minimum positive integer for which
there is a sequence of tuples (sets) with at most $i$ atoms each, $t_0, t_1, \ldots, t_n$
such that 1)~$t_0$ is true in the initial state of $P$, 2)~any shortest plan $\pi$  that achieves $t_i$ in $P$
can be extended into a shortest plan that achieves $t_{i+1}$ by extending   $\pi$ with one action, and
3)~any shortest plan that achieves $t_n$ is also a shortest plan for $P$.

One way to understand this definition is that the problem  width $w(P)$ is at most  $i$ if  there is a
``trail of  stepping stones''  $t_0, t_1, \ldots, t_n$ to reach the problem goal such that
A)~these ``stepping stones''  contain at most $i$ atoms each, B)~each  stepping stone $t_{i+1}$ is
at distance 1 from the previous one $t_i$ when $t_i$ has been reached following the ``trail'',
C)~the ``trail''  preserves ``optimality''; i.e., no  tuple $t_i$ can be reached in less than $i$ steps.

While this notion of width and the iterated  width algorithms that are based on it  have been designed
for problems where a goal state needs to be reached, the notions remain
relevant  in  optimization problems as well. Indeed, if a  good path is made of
states $s_i$ each of which has a low width, \IW\ can be made
to find such path in low polynomial time for a small value of the $k$ parameter.
Later on we will discuss  a slight  change  required in \IW\ to enforce this property.
% which is actually  not yet implemented in the code that is tested below.

\section{The Algorithms for the Atari Games}

The number of nodes \emph{generated} by $\IW(1)$ is $n \times D \times b$ in the worst case,
where $n$ is the number of variables, $D$ is the size of their domains, and $b$ is the number of
actions per  state.  This same number in breadth-first search is not linear in $n$
but exponential. For the Atari games,  $n=128$, $D=256$, and $b=18$, so that the product is
equal to $589,824$, which is large but feasible. On the other hand, the number
of nodes generated by $\IW(2)$ in the worst case is  $(n \times D)^2 \times b$, which is
equal to $1,073,741,824$ which is too large, forcing us to consider only a fraction of
such states. For classical planning problems, the growth in the number of nodes from
$\IW(1)$ to $\IW(2)$ is not that large, as the variables are boolean. Indeed, we could have
taken the state vector for the Atari games as a vector of 1024 boolean variables,
and apply these algorithms to that representation. The results would be different.
In such a case, $\IW(1)$ and $\IW(2)$  would generate $n' \times D' \times b$ and
$(n' \times D')^2 \times b$ states, which for $n'=1024$ and $D=2$
represent  $36,864$ and $7,5497,472$ states respectively.
These are feasible numbers, but we haven't used  this representation 
under the assumption that the correlations among the bits in each one of
the 128 words of the state vectors are meaningful.
In summary, from the basic \IW\ algorithm we are testing only the first
call $\IW(1)$.

\IW\ is a purely exploration algorithm that does not take into account  the accumulated
reward for selecting the states to consider. As a simple variant that combines
exploration and exploitation, we evaluated a \emph{best-first search} algorithm
with two queues: one queue  ordered first by  novelty measure
(recall that novelty one means that the state is the first one to make some atom true),
and a second queue ordered by accumulated reward. In one iteration, the best first
search picks up the best node from one queue, and in the second iteration it picks
up the best node from the other queue. This way for combining multiple heuristics
is used in the LAMA planner \cite{richter:lama}, and was introduced in the
planner Fast Downward \cite{helmert:fd}. In addition, we break ties
in the first queue favoring states with largest accumulated reward, and
in the second queue, favoring states with smallest novelty measure. Last, when
a node is expanded, it is removed from the queue, and its children are placed on
both queues. The exception are the nodes with no accumulated reward  that are
placed in the first queue only. We refer to this best-first algorithm
as \BFS.

For  the experiments below, we added two simple variations to  $\IW(1)$ and \BFS.
First, in the breadth-first search underlying $\IW(1)$, we generate the children in
random order. This makes the executions that results from the
$\IW(1)$ lookahead less susceptible to fall into loops; a potential
problem in local search algorithms with no memory or learning.
Second, a discount factor $\gamma=0.995$ is used in both algorithms
for discounting future rewards like in \UCT.
For this, each state $s$ keeps its depth $d(s)$ in the search tree,
and if state s' is the child of state $s$ and action $a$,
$R(s')$ is set to $R(s) + \gamma^d(s´) r(a,s)$.
The  discount factor  results in a slight
preference for rewards that can be reached earlier,
which is a  reasonable heuristic in  on-line settings
based on lookahead searches.

\section{Experimental Results}

\begin{table*}
\centering
% \addtolength{\tabcolsep}{-2pt}
% {\scriptsize
{\small
\begin{tabular}{@{}l|r|rr|rr|rr|rr@{}}
%% table header
% \hline\\[-3ex]
%\hline
\multicolumn{1}{c|}{} & \multicolumn{1}{c|}{Random} & \multicolumn{2}{c|}{\BRFS}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS}&
   \multicolumn{2}{c}{\UCT}
% \\[.2ex]
\\
% \hline\\\\[-3ex]
\hline
%% table body
\multicolumn{1}{c|}{Game}  &
% \multicolumn{1}{c|}{Avg. Score} &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
 \multicolumn{1}{c|}{Score} &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c}{Time}
% \\[.2ex]
\\[.2ex]
% \hline\\
\hline
\Asterix        & 210    & 3900  & 32.24 & 119000  & 17.08 & 56320  & 0.95  & \textbf{296300} (290700) & 10.82\\
\BeamRider     & 448.8 & 519.2 & 43.48 & 3000          & 6.02  & 3804        & 0.92  & \textbf{4320} (6624)   & 42.91\\
\Breakout       & 1    & 2     & 51.56 & 383.6         & 1.37  & 421.2       & 3.95  & \textbf{512.4} (364)   & 26.03\\
\Enduro          & 0   & 0.4  & 40.92 & \textbf{500}   & 38.13 & 292.8      & 13.12 & 306.8 (286)             & 32.14\\
\Freeway         & 0   & 0     & 71.84 & \textbf{29.4} & 15.19 & 19.6       & 4.44  & 1 (0)    & 90.64\\
\Pong           & -20.4 & -20.2 & 41.51 & 16.4         & 18.45 & \textbf{21} & 39.69 & \textbf{21} (21) & 38.37\\
\Qbert           & 185   & 160  & 42.83 & 1545         & 1.82  & 14725       & 1.76  & \textbf{18730} (17343) & 28.61\\
\Seaquest       & 192   & 200   & 40.01 & \textbf{4274}& 14.72 & 4070        & 1.59  & 3406 (5132)      & 26.16\\
\SpaceInvaders & 187    & 115   & 46.11 & \textbf{2670} & 12.1  & 1327       & 2.25  & 2634 (2718)  & 19.66
\end{tabular}
}

 \caption{\small Performance that results from random policy and various  lookahead algorithms in 9 Atari games.
The algorithms  are \BRFS, $\IW(1)$, \BFS, and \UCT. Averages  over 5 runs shown with ``random'' seeds
$0$ to $4$, and maximum episode duration of $18,000$  frames. Every algorithm limited to a budget of
100,000 simulated frames, except for UCT limited to 500 rollouts, 300 frames deep instead. Average scores and CPU times per action selection shown.
Performance of UCT results  reported in \cite{bellemare:jair2013} shown in parenthesis.
Numbers  in bold show best performer in terms of average score. }
\label{performance}

\end{table*}

We tested  $\IW(1)$ and \BFS\  over $9$ games: those considered in the \emph{training} set
of \cite{bellemare:jair2013}, abbreviated  as BNVB, and those discussed in \cite{deep-mind-atari}.
In Table~\ref{performance},  the performance of  $\IW(1)$ and \BFS\
% , characterized by the average score obtained over all episodes, the standard deviation
% of the score obtained and time (in seconds) per lookahead,
is compared with that of \UCT, breadth-first search (\BRFS), and random action selection.
The discount factor used by all the algorithms is $\gamma=0.995$ and the average scores
reported for \UCT\ in  BNVB  are included in  parenthesis for reference. 
Our experimental setup  follows the one in BNVB with two exceptions: rather than
considering $10$ runs per games, we considered just $5$ due to time constraints (*** ???),
and  a maximum budget of  $150,000$ simulated frames is applied to all the  algorithms, 
including  \UCT\ that is limited to  $500$ base policy rollouts of depth $300$.
% \footnote{In \UCT,  $500$ base policy rollouts
% of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
The bound on the number of simulated frames is like a bound on lookahead time, 
as most of the time in the lookahead search is spent in the calls to the emulator
for computing the next RAM state.  Also in order to save time, as in BNVB,   all of the algorithms 
reuse the subtree from  the previous lookahead that is rooted in the selected
child, deleting its siblings and their descendants. More precisely, no calls to the
emulator are done  for transitions that are cached in that subtree. As a result
the number of nodes in the lookahead tree may be larger actually than the budget.
In addition, in  \IW, the states that are ``reused''  from the  previous searches
are ignored in the computation of the novelty of new states so that more states 
can  escape pruning. Otherwise, \IW\ often doesn't use the full budget. This is not 
needed in \BFS\ which does no pruning.  $\IW(1)$ and \BFS\ have been  limited 
search up to a depth of $m=1500$  frames and up to $m=20,000$ frames per root branch.
This is to avoid the search for going too deep or being too committed to a single
root action. (**** Please revise ****).

Table~\ref{performance} shows that $\IW(1)$ and \BFS\ both outperform \BRFS\  and the random baseline.
\BRFS\ rarely collects any reward, as the depth of the \BRFS\ search tree
results in  a lookahead of $0.3$ seconds ($20$  frames or $4$ nodes deep).
On the other hand, both $\IW(1)$ and \BFS\ reach states that are  up to $350$--$1500$ frames deep
($70$--$260$ nodes or $6$--$22$ seconds), even if $\IW(1)$ doesn't always use all the simulation frames
allocated due to its agressive pruning.
%% case of \BFS, due to having a novelty of $0$ and the planning horizon being limited to $m$
%%% (*** What is novelty $0$??? Not clear why BFS doesn't take all the budget.

Interestingly, simple $\IW(1)$ outperforms \UCT\ in $4$ out of $9$ games,
both in terms of  average scores and average times per call. \BFS\ is
faster than $\IW(1)$ and an order-of-magnitude faster than \UCT\ in general,
and it  outscores \UCT\ in $2$ of the games (\Freeway\ and
\Seaquest). Compared with the \UCT\ results in   \cite{bellemare:jair2013}, shown in parenthesis,
both $\IW(1)$ and \BFS\  outperform \UCT\ in $3$ of the  $9$ games (\Breakout, \Enduro\ and \Freeway),
with \BFS\ and \UCT\ tied in  \Pong. It'd  be very interesting to see the results of
\BFS\ when given more time per action, but we don't have those results yet.
As mentioned in the previous section, the order in which $\IW(1)$ and \BFS\ generate the children
of an expanded node are randomized, something which has a potential effect on the  novelty measures
of nodes, and hence in the search graph. We found that this randomization helps, in particular
in \Qbert, where not randomizing results  in $\IW(1)$ being trapped into  loops sometimes.

We also  note that in \Freeway,  both $\IW(1)$ and \BFS\ attain a score similar to the
baseline semi-random algorithm \emph{Perturb} in  \cite{bellemare:jair2013},
which outperforms   both \UCT\ and  the reinforcement learning methods.
Likewise,  $\IW(1)$ outperforms the best learning algorithm in \cite{deep-mind-atari}
in $4$ out of the  $7$ games considered, and \BFS\ does so in $5$  of
the $7$ games.

We do not report the performance of $\IW(2)$ because  in our
preliminary tests and according to the discussion in the previous section,
it doesn't  appear to improve much on  \BRFS, even if  it results in a lookahead
that is $5$ times deeper,  still far too shallow  to compete with the other
planning algorithms.

\section{Exploration and Exploitation}

The notion of width underlying the iterated width algorithm was developed in the context
of classical planning in order to understand why most of the hundreds of existing benchmarks
appear to be relatively simple for current planners, even
though classical planning is PSPACE-complete \cite{bylander:complexity}. A  partial answer is that
most of these domains have a low width, and hence, can be solved in low polynomial time (by \IW)
when goals contain a single atom.  Benchmark problems with multiple atomic goals tend to be
easy too, as the goals can often be achieved one at a time after a simple goal ordering \cite{nir:ecai2012}.

In the iterated width algorithm, the key notion is the \emph{novelty measure} of a state in the underlying
breadth-first search. These novelty measures make use of the factored representation of the states for
providing a structure to the search: states that have width 1 are explored first in linear time,
then states that have  width 2  are explored in quadratic time, and so on. In classical planning problems
with atomic goals, this way of organizing the search  pays off both theoretically and practically.

The use of ``novelty measures'' for guiding an optimization search while ignoring the function that is being
optimized is   common to the  novelty-based search methods developed independently
in the context of genetic algorithms \cite{lehman:novelty}. %%% bibtex for this below
In these methods individuals in the population are not ranked according to the  optimization function
but in terms of how ``novel'' they are in relation to the rest of the population, thus encouraging diversity
and exploration rather than (greedy) exploitation. The actual definition of novelty
in such a case is domain-dependent; for example, in the evolution of a controller
for guiding a robot in a maze, an individual controller  will not be ranked by how
close it takes the robot to the goal (the greedy measure), but by the distance between
the locations that are  reachable with it, and the locations reachable with
the  other controllers in the population (a diversity measure). The novelty measure used by \IW, on the other hand, is domain-independent and it
is determined by the structure of the states as captured by the problem variables.

The balance between exploration and exploitation has received considerable attention in
reinforcement learning \cite{sutton:book}, where both are required for converging to an
optimal behavior. In the Atari games, as in other deterministic problems, however, ``exploration''
is not needed for optimality purposes, but just for improving the effectiveness of the
 lookahead search. Indeed, a best-first search algorithm guided only  by (discounted)
accumulated reward will deliver eventually  best moves, but it will not be as effective
over small time windows, where like breadth-first search it's likely not to find any rewards
at all. The \UCT\ algorithm provides a method for balancing exploration and exploitation,
which  is  effective over small time windows. The \BFS\ algorithm above with two queues
that alternate, one guided by the novelty measures  and the other by the accumulated reward,
provides a different scheme. The first converges to the optimal behavior asymptotically; the
second in a bounded number of steps, with the caveat below.

% Alternatively, novelty-guided probes, possibly with
% some noise added, could potentially pay off as  a  base policy for \UCT\ in replacement
% of random roll outs, as they apparently manage to exploit the structure of many problems.

\section{Duplicates and Optimality}

The notions of width and the \IW\  algorithm guarantee that  states with low width will be generated
in low polynomial time through \emph{shortest} paths. In the presence of rewards like the Atari games,
however, the  interest is not  in the shortest paths but in the \emph{best} paths; i.e, the paths with
maximum reward.  \IW\ may actually  fail to find such paths even when calling \IW($k$) with a high 
$k$ parameter. Optimality  could be  achieved then  by replacing the breadth-first search
underlying \IW($k$) by Dijkstra's algorithm yet such a move would make the relation between
\IW\ and the notion of width less clear. Another option is to change  \IW\ to comply with 
a different property; namely, that if there is a ``rewarding'' path made up of states of 
low width, then \IW\ will find such paths or better ones in time that is exponential 
in their width. For this actually a simple change in \IW\ suffices: when generating a state 
$s$ that is a \emph{duplicate} of a state $s'$ that has been previously generated and not pruned, 
$s'$ is replaced by $s$ if $R(s) > R(s')$. Moreover in such a case, the change in reward 
must be propagated to the descendants of $R(s')$ that are in memory. This is similar to the change required 
in A* for preserving optimality when moving from consistent to inconsistent heuristics \cite{pearl:heuristics}. 
The alternative is to ``reopen'' such nodes.  The same change is actually in \BFS\ to ensure that, 
if given enough time, \BFS\ will actually find the optimal paths. The code used for \IW\ and \BFS\ in the
experiments above does not implement this change, however, as the overhead involved in checking for duplicates 
in some test cases did not appear to pay off. More experimentation is probably needed.

\section{Summary and Future Work}

We have shown experimentally that width-based algorithms like $\IW(1)$ and \BFS\
that originate in work in classical planning,  can be used to play the Atari video
games and to achieve  state-of-the-art performance. This is the result of a 
structured exploration of the state that manages to combine the scope of blind
search algorithms with the efficiency of heuristic search methods. 
%
\Omit{
We are currently addressing three
issues, which we haven't managed to get in time for this submission. First, we want to extend the
evaluation to all the games featured in \cite{bellemare:jair2013}, and to add more runs per  game.
Second, we want to evaluate the various algorithms for different fixed time windows for decisions.
In the table above, for example, \UCT\ is often taking an order-of-magnitude more time per decision
than  \BFS. Third, we need to revise the way duplicate states are handled in \IW\ and \BFS\ to ensure
that the best paths among the states that have been expanded are returned.Finally,}
%
There are many issues worth studying, we just mention two of them.
One problem of all planning methods in Atari  games is that by using the 
RAM state of the Atari console they ``cheat'' a bit, as they use information that
is not available to the human player. In addition, a limitation of our
algorithms is that they do not get better with experience.  
In the future, we would like to explore whether  variations of these algorithms 
can be used to play  from the information in the  screen state as opposed to the RAM state, 
and the potential uses of the  ideas underlying these algorithms for learning.

%%% - Learning in same instance, across diff instances of same game, etc.
%%%% - Learning controllers vs. on-line planning, etc

%%% - Learning in same instance, across diff instances of same game, etc.
%%%% - Learning controllers vs. on-line planning, etc

\bibliographystyle{aaai}
\bibliography{control}

\end{document}